---
title: "Data Camp Class Assignment"
subtitle: "Ollie Murphy"
author: "working with Hannah Korte"
date: "9/11/2021"
output: html_document
---

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)
a = "Ollie"
b1 = 45.6
b2 = "45.6"
c1 = 0:3
```
Q1: What type of data is contained in the variable a

- character

Q2: What type of data is contained in the variable b1

- numeric

Q3: What type of data is contained in the variable b2

- character

Q4: What happens when you add b1 and b2 and why?

- Error in b1 + b2 : non-numeric argument to binary operator. The variables cannot be added because R cannot do arithmetic with character variables.

Q5: Are the variables b1 and c1 of the same type?

- no

Q6: Explain what happens when you add b1 and c1. Consider both the number of elements in each variable and the data types.

- when you add b1 and c1, R adds the numeric value of b1 (45.6) to each number in the c1 vector, resulting in 45.6 46.6 47.6 48.6

Q7: Show the R code you used to create v1.
```{r echo=TRUE}
v1 = c(-2:2)
v1
```
Q8: Show the R code you used to create v2.
```{r echo=TRUE}
v2 = v1 * 3
v2
```
Q9: Show the R code you used to calculate the sum of elements in v2.
```{r echo=TRUE}
sum(v2)
```
Q10: Show the code you used to create mat_1.
```{r}
vec_4 = 1:12
mat_1 = matrix(vec_4, nrow = 3, ncol = 4, byrow = TRUE)
mat_1
```
Q11: Show the code you used to create mat_2.
```{r}
mat_2 = matrix(vec_4, nrow = 3, ncol = 4, byrow = FALSE)
mat_2
```
Q12: Show the R code you used to create my_list_1.
```{r}
my_list_1 = list("two" = 5.2, "one" = "five point two", "three" = c(0:5))
```
Q13: Show the R code that would select the third element of the list.
```{r}
my_list_1[3]
```
Q14: Show the R code that selects the list element with the name “one”.
```{r}
my_list_1$one
my_list_1[["one"]]
```
Q15: Show the R code that you used to create my_bool_vec.
```{r}
my_vec = rep(1:3, 5)
my_bool_vec = c(my_vec == 3)
data.frame(my_vec, my_bool_vec)
```
Q16: Show the R code that you used to perform the subsetting.
```{r}
my_vec[my_bool_vec]
```

