---
title: "Lab 10"
author: "Ollie Murphy"
date: "11/23/2021"
output: 
  html_document:
    toc: TRUE
    toc_float: TRUE
    toc_collapsed: TRUE
    theme: "lumen"
---

# ANOVA by hand

### 1. Submit the code you used to build your ANOVA by hand.

```{r, results = FALSE}
rm(list = ls())

require(here)
rope = read.csv(here("data", "rope.csv"))
rope$rope.type = factor(rope$rope.type)
                        
n_obs = nrow(rope)
n_groups = length(unique(rope$rope.type))
                        
ss_tot = sum((mean(rope$p.cut) - c(rope$p.cut))^2)
df_tot = (n_obs - 1)
                        
agg_sq_resids =
 aggregate(
 x = rope$p.cut,
 by = list(rope$rope.type), 
 FUN = function(x) sum((mean(x) - c(x))^2))
                    
ss_within = sum(agg_sq_resids$x)
df_within = (n_obs - n_groups)
                        
ss_among = (ss_tot - ss_within)
df_among = (n_groups - 1)
                        
ms_within = ss_within / (n_obs - n_groups)
ms_among  = ss_among / (n_groups - 1)
                        
f_ratio = ms_among / ms_within
f_pval = pf(f_ratio, df_among, df_within, lower.tail = FALSE)

```

# Model Assumptions

### 2. Examine the conditional boxplot in the Partitioning Variance: Within-Group section of the walkthrough. Based on the figure, do you think there are equal variances among the groups?

- Based on the group conditional boxplot in Partitioning Variance: Within-Group, I do not this there are equal variances among the group. I concluded this based on the widely-varying heights of the boxes for each rope type.

### 3. Conduct a Bartlett test to assess the homogeneity of variances of the percent cut among the rope type groups. Report the p-value.

```{r}
bartlett.test(p.cut ~ rope.type, data = rope)
```
- p = 0.00143

### 4. Given your graphical assessment (question 2) and the Bartlett test, do you think an ANOVA-type analysis is appropriate on the raw data? Explain why or why not.

- The low p value of the Bartlett test, which indicates that I can discard the null hypothesis of equal variance among the groups, agrees with my inital assessment of variation in the data set: it is nor equal among rope types. Because ANOVA is a group 1 model, which have an assumption of homoskedasicity (constant variance), ANOVA is likely not appropriate for this data set.

# Model Coefficients and Group Means

```{r}
fit_rope_1 = lm(p.cut ~ rope.type, data = rope)
summary(fit_rope_1)
```

### 5. Which rope type is the base case?

- The "BLAZE" rope type is the base case in this table.

### 6. What is the mean percent cut of the base case rope? Show your calculation using value(s) from the model coefficient table.

- The mean percent cut of the BLAZE rope is 36.7 %. This did not require calculation, as the intercept coefficient is interpreted as the intercept mean.

### 7. What is the mean percent cut rope type XTC? Show your calculation using value(s) from the model coefficient table.

- The mean percent cut of the XTC rope is 26.5%. I calculated this by taking the intercept coefficient (36.7) and adding the XTC slope coefficient of -10.2%, which gave me 26.5%.
